---
title: 'Assignment 10: Data Scraping'
author: "Jiyeong Pyo"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Rename this file `<FirstLast>_A10_DataScraping.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure your code is tidy; use line breaks to ensure your code fits in the knitted output.
5. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up 
1. Set up your session:

* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Check your working directory

```{r 1. Project Setup, message = FALSE}
#1 

library(tidyverse)
library(lubridate)
library(here); here()
library(rvest)

```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2024 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2024>
 
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website}
#2 
webpage <- read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2024')
webpage
```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PWSID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Day Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values (represented as strings)".

```{r scrape.the.data}
#3 
water_system_name <- webpage %>% 
  html_nodes('div+ table tr:nth-child(1) td:nth-child(2)') %>% 
  html_text()
water_system_name

PWSID <- webpage %>% 
  html_nodes('td tr:nth-child(1) td:nth-child(5)') %>% 
  html_text()
PWSID

Ownership <- webpage %>% 
  html_nodes('div+ table tr:nth-child(2) td:nth-child(4)') %>% 
  html_text()
Ownership

MGD <- webpage %>% 
  html_nodes('th~ td+ td') %>% 
  html_text()%>%
  as.numeric()
MGD

```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in chronological order. You can overcome this by creating a month column manually assigning values in the order the data are scraped: "Jan", "May", "Sept", "Feb", etc... 

5. Create a line plot of the maximum daily withdrawals across the months for 2024, making sure, the months are presented in proper sequence.

```{r create.a.dataframe.from.scraped.data}
#4 

#Create a dataframe of withdrawals
df_withdrawals <- data.frame("Month" = c(1,5,9,2,6,10,3,7,11,4,8,12),
                             "Year" = rep(2024,12))

df_withdrawals <- df_withdrawals %>% 
  mutate(water_system_name= !!water_system_name,
         PWSID = !!PWSID,
         Ownership = !!Ownership,
         MGD = !!MGD,
         Date = my(paste(Month,"-",Year)))
#5 
#Plot 
ggplot(df_withdrawals,aes(x=Date,y=MGD)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  labs(title = paste("Maximum daily withdrawals in 2024 in",water_system_name),
       y="Withdrawal (mgd)",
       x="Date")

```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function with two inputs - "PWSID" and "year" - that:
  - Creates a URL pointing to the LWSP for that PWSID for the given year
  - Creates a website object and scrapes the data from that object (just as you did above)
  - Constructs a dataframe from the scraped data, mostly as you did above, but includes the PWSID and year provided as function inputs in the dataframe. 
  - Returns the dataframe as the function's output


```{r construct.a.scraping.function}
#6. 

scrape.it <- function(PWSID, Year){
  
  #Retrieve the website contents 
  the_website <- read_html(paste0('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid='
                                  ,PWSID,'&year=', Year))
  
  #Set the element address variables 
  water_system_name_tag <- 'div+ table tr:nth-child(1) td:nth-child(2)'
  PWSID_tag <- 'td tr:nth-child(1) td:nth-child(5)'
  Ownership_tag <- 'div+ table tr:nth-child(2) td:nth-child(4)'
  MGD_tag <- 'th~ td+ td'
  
  #Scrape the data items
  water_system_name <- the_website %>% html_nodes(water_system_name_tag) %>% html_text()
  PWSID<- the_website %>%   html_nodes(PWSID_tag) %>%  html_text()
  Ownership <- the_website %>% html_nodes(Ownership_tag) %>% html_text()
  MGD <- the_website %>% html_nodes(MGD_tag) %>% html_text()
  
  #Convert to a dataframe
  df_withdrawals_2 <- data.frame("Month" = c(1,5,9,2,6,10,3,7,11,4,8,12), 
                               "Year" = rep(Year,12),
                               "Max Day Use" = as.numeric(MGD)) %>% 
    mutate(water_system_name = !!water_system_name,
           PWSID = !!PWSID,
           Ownership = !!Ownership,
           Date = my(paste(Month,"-",Year)))
  
  #Return the dataframe
  return(df_withdrawals_2)
}
```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2020. Then show the structure of the dataframe with either the `str()` or the `glimpse()` function. 

```{r fetch.and.plot.Durham.2020.data}
#7 
Year = 2020
PWSID= '03-32-010'

the_dfs <- map(Year,scrape.it,PWSID=PWSID)
df_durham <- bind_rows(the_dfs)
str(df_durham)

ggplot(df_durham,aes(x=Date,y=`Max.Day.Use`)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  labs(
    title = paste(
      "Max Daily Withdrawals for", water_system_name,"in",Year),
    y="Withdrawal (mgd)",
    x="Date"
    )

```

8. Use the function above to extract data for Cary (PWSID = '03-92-020') in 2020. Combine this data with the Durham data collected above and create a plot that compares Cary's to Durham's water withdrawals.

```{r fetch.and.plot.Cary.2020.data}
#8 
#Extract for Cary
Year = 2020
PWSID= '03-92-020'

cary <- map(Year,scrape.it,PWSID=PWSID)
df_cary <- bind_rows(cary)
str(df_cary)

#combine df
df_cary_durham  <- bind_rows(df_cary,df_durham)

#Plot
ggplot(df_cary_durham,aes(x=Date,y=`Max.Day.Use`,color=water_system_name)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  labs(title =paste("Cary vs Durham: Max Daily Withdrawals in",Year),
       y="Withdrawal (mgd)",
       x="Date")

```


9. Use the code & function you created above to plot Cary's max daily withdrawal by months for the years 2018 thru 2023.Add a smoothed line to the plot (method = 'loess').

> TIP: See Section 3.2 in the "10_Data_Scraping.Rmd" where we apply "map2()" to iteratively run a function over two inputs. Pipe the output of the map2() function to `bind_rows()` to combine the dataframes into a single one, and use that to construct your plot.

```{r fetch.and.plot.Cary.multiyear.data }
#9 
#Extract for Cary
Year = c(2018:2023)
PWSID= '03-92-020'

cary <- map2(PWSID, Year, scrape.it) %>% 
  bind_rows()
str(cary)

#Plot
ggplot(cary,aes(x=Date,y=`Max.Day.Use`)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  scale_x_date(date_breaks = "1 year",date_labels = "%Y")+
  labs(title = paste("Max Daily Withdrawals in Cary through",
       Year[1],"-",Year[length(Year)]),
       y="Withdrawal (mgd)",
       x="Year")


```

Question: Just by looking at the plot (i.e. not running statistics), does Cary have a trend in water usage over time?
<br>
> Answer: It seems that Cary do have a strong seasonality in water usage over time. Cary's water withdrawals surge in summer season while there is no distinctive long term trend.
>
