---
title: "Lab2 Jiyeong Yui"
author: "Jiyeong Pyo, Yui Kikuchi"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 2
---

```{r chunk_1, include=F}

library(moments) #package for skewness
library(knitr) #package for making tables (kable)
library(tidyverse) #multiple packages for data wrangling
library(gt) # a package to make tables
library(lubridate) # a package to manipulate dates
library(tidyr)
library(janitor)
library(here)

```


```{r chunk_2,include=F}
PEC <- read.csv(
  file=here("~/EDE_Fall2025/Justina/ENV710/PEC_final.csv"), 
  stringsAsFactors = TRUE
   )
PEC_remove <- PEC %>%
 # mutate(date=ymd(`month_year`))%>% #making date variable
  group_by(canton,year)%>%
  count()%>%
  filter(n<12)

PEC_filtered <-PEC %>% 
  filter(!((canton=="Chilla" & year==2015) | (canton=="Chilla" & year==2020)|(canton=="Esmeraldas" & year==2020)|(canton=="Olmedo2" & year==2018)))

```

In terms of cleaning the Data, Chilla's first year(2015) and the last year(2020) was missing and Esmeraldas's last year(2020) data was missing. So that the comparison will be excluding these canton.  


[B]
```{r chunk_3, Filtering PEC high canton,include=F}
PEC_High <- PEC_filtered %>%
  subset(PEC_filtered$high_adopt==1)
glimpse(PEC_High)
```
```{r}
ea.summary<-PEC_High |>  
  
  summarise(n = length(na.omit(kwh_total)),
            min=min(kwh_total, na.rm=TRUE),  
            max=max(kwh_total, na.rm=TRUE),  
            mean=mean (kwh_total, na.rm=TRUE),
            stdev=sd(kwh_total, na.rm=TRUE),
            q25=quantile(kwh_total, 0.25, na.rm=TRUE),
            q75=quantile(kwh_total, 0.75, na.rm=TRUE),
            skewness=skewness(kwh_total, na.rm=TRUE))


ea.summary 
```
```{r,include=F}
PEC_High_filtered <- PEC_High %>%
  group_by(canton,year)%>%
  summarise(kwh_total=sum(kwh_total))%>%
  filter(year=='2015'|year=='2020')
  
  
head(PEC_High_filtered)
```
```{r,message=FALSE,include=F}
PEC_High_filtered<-PEC_High_filtered %>%
  mutate(year=factor(year, levels=c("2015","2020")))

p<-ggplot(data=PEC_High_filtered, aes(x=kwh_total, fill=year))+
geom_histogram(col="black")+
scale_fill_manual(values=c("royalblue", "gray"))

p + facet_grid(year ~ .)+
labs(x="Energy consumption (kWh)", title="Comaprison of energy consumption in high adopted area",
     fill="year") +
 theme(
    plot.title = element_text(size = 8, color = "navy", hjust = 0.5),
    axis.title.x = element_text(size = 8, color = "black"),
    axis.title.y = element_text(size = 8, color = "black")
  )
```

```{r,filtering for wide,message=FALSE,include=F}
PEC_High_filtered_wide<-PEC_High_filtered %>%
  dplyr::select(canton, year, kwh_total)%>%
  pivot_wider(names_from = year, values_from = kwh_total) %>%
  rename(first_year="2015")%>%
  rename(last_year="2020")

PEC_High_filtered_wide
```
3. Statistical Analysis and Discussion

a. Comparison of Means
```{r chunk_6, Paired t-test,echo=FALSE,include=F}
t.test(PEC_High_filtered_wide$first_year, PEC_High_filtered_wide$last_year,alternative="greater", paired=T)
```

For comparing the mean electricity consumption levels between first year(2015) and last year(2020)in high adoption areas, we used paired t-test methods.sfAccording to the high p-value((t = -2.4966, df = 54, p = 0.9922),we cannot reject the null hypothesis which means that there is no statistically significant increase in electricity consumption. Moreover,the difference in electricity consumption was calculated as 2020 minus 2015 and the negative mean difference(-8416716) indicates that there was no consumption increase over time.


```{r,echo=FALSE,include=F}
PEC_High_filtered_wide<-PEC_High_filtered_wide  |>
      mutate(diff=last_year-first_year)

head(PEC_High_filtered_wide)
```

The distribution of the differences shows that the electricity consumption between 2020 and 2015 in high adoption area is highly right-skewed,indicating the majority experienced little or no consumption change while few cantons experienced consumption increase.

```{r,message=FALSE,include=F}
ggplot(data = PEC_High_filtered_wide, aes(x = diff)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 100) +
  labs(
    title = "Electricity Consumption Levels in High Adoption Areas",
    subtitle = "Difference between 2015 and 2020",
    x = "Difference (2020 - 2015)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 8, color = "navy"))
```
b. Discussion of assumptions
We conducted Shaprio-Wilk test for normality test and it showed strong deviation from normality indicating the t-test assumption of normality is violated (W = 0.26874, p-value = 4.638e-15).

```{r,message=FALSE,include=F}
#Null hypothesis: the sample in our case the differences was pulled from a normally distributed population. 
shapiro.test(PEC_High_filtered_wide$diff)
skewness(PEC_High_filtered_wide$diff)
```


```{r,checking Logarithmic,message=FALSE}
PEC_High_filtered_wide<-PEC_High_filtered_wide|>
    mutate(log.first_year=log(first_year),    				       	
           log.last_year=log(last_year), 
			     diff.in.log=(log.first_year-log.last_year),
           log.ratio=log(first_year/last_year))
head(PEC_High_filtered_wide, 6)
```
For correcting skewness, we used log-transformation. After transformation, a paired t-test on the logged data showed no significant increase even after the transformation(t = -11.841, df = 54, p-value = 1). 

```{r,t-test of logged data,message=FALSE,include=F}
t.test(PEC_High_filtered_wide$log.first_year, PEC_High_filtered_wide$log.last_year, paired=TRUE, alternative = "greater")
```
The Shapiro-Wilk test indicates the normality is not violated after log transformation(W = 0.96951, p-value = 0.1751).
```{r,shapiro test,message=FALSE,include=F}
shapiro.test(PEC_High_filtered_wide$diff.in.log)
skewness(PEC_High_filtered_wide$diff.in.log)
```
 
c. Discussin of appropriate test
  In conclusion, the raw data violated the normality condition. After log transformation, the normality condition is satisfied but there was no consumption increase.So for this comparison, the log-transformed data were considered the most appropriate.
  

[C]

```{r,filtering,message=FALSE,include=F}
PEC_Low <- PEC_filtered %>%
  subset(PEC_filtered$high_adopt==0)
glimpse(PEC_Low)
```

```{r,message=FALSE,include=F}
ea.summary<-PEC_Low |>  
  
  summarise(n = length(na.omit(kwh_total)),
            min=min(kwh_total, na.rm=TRUE),  
            max=max(kwh_total, na.rm=TRUE),  
            mean=mean (kwh_total, na.rm=TRUE),
            stdev=sd(kwh_total, na.rm=TRUE),
            q25=quantile(kwh_total, 0.25, na.rm=TRUE),
            q75=quantile(kwh_total, 0.75, na.rm=TRUE),
            skewness=skewness(kwh_total, na.rm=TRUE))


ea.summary 
```



```{r, filtering by years,message=FALSE,include=F}
PEC_Low_filtered <- PEC_Low %>%
  group_by(canton,year)%>%
  summarise(kwh_total=sum(kwh_total))%>%
  filter(year=='2015'|year=='2020')
  
  
head(PEC_Low_filtered)
```

```{r,echo = FALSE,include=F}
PEC_Low_filtered<-PEC_Low_filtered %>%
  mutate(year=factor(year, levels=c("2015","2020")))

p<-ggplot(data=PEC_Low_filtered, aes(x=kwh_total, fill=year))+
geom_histogram(col="black")+
scale_fill_manual(values=c("royalblue", "gray"))

p + facet_grid(year ~ .)+
labs(x="Energy consumption (kWh)", title="Comaprison of energy consumption in low adopted area",
     fill="year") +
 theme(
    plot.title = element_text(size = 8, color = "navy", hjust = 0.5),
    axis.title.x = element_text(size = 8, color = "black"),
    axis.title.y = element_text(size = 8, color = "black")
  )
```


```{r,message=FALSE,include=F}
PEC_Low_filtered_wide<-PEC_Low_filtered %>%
  dplyr::select(canton, year, kwh_total)%>%
  pivot_wider(names_from = year, values_from = kwh_total) %>%
  rename(first_year="2015")%>%
  rename(last_year="2020")

PEC_Low_filtered_wide
```
 a. Comparison of means results

```{r chunk, Paired t-test,message=FALSE,include=F}
t.test(PEC_Low_filtered_wide$first_year, PEC_Low_filtered_wide$last_year,alternative="greater", paired=T)
```

The paired t-test in low adoptation areas suggested no significant increase(t = -2.7019, df = 52, p-value = 0.9954). 
```{r,message=FALSE,include=F}
PEC_Low_filtered_wide<-PEC_Low_filtered_wide  |>
      mutate(diff=last_year-first_year)

head(PEC_Low_filtered_wide)
```
b. Discussion of assumptions
```{r,echo = FALSE,include=F}
ggplot(PEC_Low_filtered_wide, aes(x = diff)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 100) +
  labs(
    title = "Electricity Consumption Levels in low Adoption Areas",
    subtitle = "Difference between 2015 and 2020",
    x = "Difference (2020 - 2015)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 8, color = "navy"))
```
The difference between in low adoption areas showed right-skewed distribution, the majority had little consumption chage and only small number of cantons showed consumption increase. It indicates the data are non-normally distributed.

```{r,message=FALSE,include=F}
shapiro.test(PEC_Low_filtered_wide$diff)
skewness(PEC_Low_filtered_wide$diff)
```
The Shapiro-Wilk test showed severe non-normality(W = 0.42532, p-value = 4.229e-13).
```{r,logged values,message=FALSE,include=F}
PEC_Low_filtered_wide<-PEC_Low_filtered_wide|>
    mutate(log.first_year=log(first_year),    				       	
           log.last_year=log(last_year), 
			     diff.in.log=(log.first_year-log.last_year),
           log.ratio=log(first_year/last_year))
head(PEC_Low_filtered_wide, 6)
```
After log transformation, the paired t-test still indicated we cannot reject the null hypothesis(t = -4.2235, df = 52, p-value = 1). The mean difference was negative value, suggesting the electricity consumption rather decreased. 
```{r,echo=FALSE,include=F}
t.test(PEC_Low_filtered_wide$log.first_year, PEC_Low_filtered_wide$log.last_year, paired=TRUE, alternative = "greater")
```
The Shaprio-Wilk test after log transformation showed
the data still violated normality(W = 0.68309, p-value = 2.054e-09), suggesting that nonparametric method needed for robustness. 


```{r,echo=FALSE,include=F}
shapiro.test(PEC_Low_filtered_wide$diff.in.log)
skewness(PEC_Low_filtered_wide$diff.in.log)
```
c. Discussion of appropriate test
 Because the Shapiro-Wilk showed non-normal even after log transformation, we conducted Wilcoxon test and the result indicated no significant shift (V = 130, p-value = 1).
```{r,message=FALSE,echo=FALSE}
wilcox.test(PEC_Low_filtered_wide$first_year, PEC_Low_filtered_wide$last_year, alternative="greater", paired = TRUE)
```


4. Conclusion

a. Scope of inference
 Since the data was measured across time, it could be not randomly sampled and limited sample size.It cannot be generalized to all regions or countries. 
 
b. Limitations and weaknesses
 Cantons are geographically close with each other so might influenced by weather condition and income levels which means there was a chance to be not independent units. 
 In addition, the data were collected over time in the same cantons so the observation could be correlated with the another year.
 
c. Real-World implications
  Both high and low adoption areas showed no statistically meaningful increase in electricity consumption.This points out the PEC program could promote electrification without increasing energy consumption. 
  However, selection bias caused by different socioeconomic should be handled. According to the paper, higher baseline electricity consumers were more likely to enroll(Gould et al., 2023). So the future policy makers should examine these bias before applying policy.

