---
title: "ENV710.002  Applied Statistical Modeling Midterm Exam"
author: "Jiyeong Pyo"
subtitle: Due 27 October 2025 at 11:45 ET
output:
  html_document:
    df_print: paged
---

#### **Duke Community Standard**
Duke University is a community dedicated to scholarship, leadership, and service and to the principles of honesty, fairness, respect, and accountability. Citizens of this community commit to reflect upon and uphold these principles in all academic and nonacademic endeavors, and to protect and promote a culture of integrity.

#### **The Pledge**
##### Students affirm their commitment to uphold the values of the Duke University community by signing a pledge that states:

####  **To uphold the Duke Community Standard:**
+ I will not lie, cheat, or steal in my academic endeavors; 
+ I will conduct myself responsibly in all my endeavors; and 
+ I will act if the Standard is compromised. 

#### I have adhered to the Duke Community Standard in completing this assignment.


#### Type your name below to state that you adhered to the Duke Community Standard while completing this exam.

<br>
Jiyeong Pyo
<br>



**1. Histograms (20 points)**
```{r libraries,include=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(broom)
library(ggthemes)
```


```{r import data,include=FALSE}
yale<-read_csv("Justina/ENV710/midterm.csv")
FEMA <-read_csv("Justina/ENV710/FEMA_regions_states.csv")
```

```{r histograms, echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
ggplot()+
  geom_histogram(yale,mapping=aes(x=gwvoteimp),bins=50,color="white",fill="lightblue")+
  theme_economist_white() +
  labs(title="Distribution of Views on Global Warming are Important",
       subtitle = "% saying a candidate’s stance on global warming matters to their vote",
       x = "Global Warming Importance")+
  theme(plot.title = element_text(size = 13, face = "bold", color = "Black"),
        plot.subtitle = element_text(size = 10, color = "black"))

ggplot()+
  geom_histogram(yale,mapping=aes(x=RISK_SCORE),bins=50,color="white",fill="lightblue")+
  theme_economist_white() +
  labs(title="Distribution of Risk score",
       subtitle = "Composite score (1–100) of county’s exposure and vulnerability to natural hazards",
       x = "Risk Score")+
  theme(plot.title = element_text(size = 13, face = "bold", color = "Black"),
        plot.subtitle = element_text(size = 11, color = "black"))
```


**2. Summary statistics (10 points)**
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required libraries
library(tidyverse)   # for data manipulation and cleaning
library(moments)     # for calculating skewness
library(gt)

# Define units for each variable
units <- tibble(
  variable = c("RISK_SCORE", "gwvoteimp"),
  units = c("Scale: 1-100", "%")
)

# Create summary statistics table
summary_table <- yale %>%
  pivot_longer(cols = c(RISK_SCORE, gwvoteimp),   # convert selected variables from wide to long format
               names_to = "variable",        # name column will store the variable names
               values_to = "value") %>%      # value column will store the variable values
  group_by(variable) %>%                     # group the data by variable name
  summarise(
    N = sum(!is.na(value)),                  # calculate the number of non-missing observations
    Mean = round(mean(value, na.rm = TRUE), 2),
    Median = round(median(value, na.rm = TRUE), 2),
    SD = round(sd(value, na.rm = TRUE), 2),
    Skewness = round(skewness(value, na.rm = TRUE), 2)
  ) %>%
  left_join(units, by = "variable") %>%      # merge in the unit information
  select(variable, units, N, Mean, Median, SD, Skewness)  # arrange columns in the desired order

# Print the summary table
gt_table <- gt(summary_table,caption="Summary Statistics",auto_align = FALSE)
gt_table
```
<br>
**3. Scatterplots (10 points) **
<br>
```{r,echo=FALSE, message=FALSE,warning=FALSE}
options(scipen = 999) 

p1 <- yale %>% 
  ggplot(aes(x = RISK_SCORE, y = gwvoteimp, label = COUNTY))+
       geom_point(size = 1) + 
       theme_minimal()+
       labs(title="Global Warming Importance ~ Risk Score",
       x = "Risk Score",
       y= "Global Warming Importance",
       caption = "*Global Warming Importance(y-axis) displays gwvoteimp variable")+
       ylim(0,100)+
       geom_text(aes(label = COUNTY), size = 3, vjust=-0.5) 
p1
```
<br>
The scatter plot shows that there is slight positive relationship between risk score and global warming importance.
The data is evenly distributed so we don't need to do additional transformation.
<br>
**4. Hypotheses and Model (30 points)**
<br>
The null hypothesis: β₁= 0 : There is no significant difference between risk score and global warming importance.
The alternative hypothesis: β₁≠ 0 : There is significant difference between risk score and global warming importance.

```{r echo=FALSE, message=FALSE, warning=FALSE}
linear <- lm(gwvoteimp ~ RISK_SCORE , data = yale)
summary(linear)
```
The simple linear regression of risk score on global warming importance shows we reject the null hypothesis,suggesting the risk score and the warming importance are correlated.(P<0.001) The test of significance (a t-test) on the coefficient tests whether the slope of the line (of the populations) is 0.07 which indicates that the 1 unit(index) increase of risk score increases 1 unit(%) of global warming importance. The R-squared value is 14% which means the percent of the variation in the global warming importance variable that is explained by the risk score.  

<br>
**5. Model Assumptions (20 points)**
<br>
```{r diagnostic plots, echo=FALSE, message=FALSE, warning=FALSE}
plot(linear,1)
```
<br>

The assumptions of simple linear regression model are the data is linear and normally distributed. The first assumption, zero conditional mean is met because there is no clear pattern across fitted values. The second assumption, variables are independent and identically distributed is met and there is no extreme outliers.

<br>
**6. Visualization: Risk across US regions (20 points) **
<br>
The Federal Emergency Management Agency(FEMA) consists of ten regions in the continental United States and territories.The risk score is varied by region.According to the violin plot, each region has different distribution. Especially, region 9 shows that the median of the data is located in the upper part and region 3,8 indicates that the median is located in the lower part. which means that the risk score can be affected by different confounding factors. 

```{r mingle the data, include=FALSE}
yale_region<-yale%>%
  dplyr::select(RISK_SCORE, STATE, STATEABBRV)
risk_by_region <- full_join(yale_region,FEMA , by = c("STATE" = "State","STATEABBRV"="Abbreviation"))
risk_by_region <- rename(risk_by_region, region= `FEMA Region`)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
Plot_FEMA<-ggplot(data=risk_by_region, aes(x=factor(region), y=RISK_SCORE,color=factor(region)))+
  geom_violin()+
labs(x="FEMA region", y="Risk Score", title="Risk Score by FEMA region") 
Plot_FEMA
```

<br>
**7. Summary statistics table of risk across regions (20 points)**
<br>
The summary table shows that the mean scores vary from 40.83 to 84.29. 4 of the 10 regions exhibit positive skewness, indicating that high-risk outliers are affecting the mean in those areas. Furthermore, the large variation in sample sizes (N) across regions suggests that the reliability of these statistics can be undermined.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Create summary statistics table
summary_table <- risk_by_region %>%
  group_by(region) %>%                     # group the data by variable name
  summarise(
    N = sum(!is.na(RISK_SCORE)),                  # calculate the number of non-missing observations
    Mean = round(mean(RISK_SCORE, na.rm = TRUE), 2),
    Median = round(median(RISK_SCORE, na.rm = TRUE), 2),
    SD = round(sd(RISK_SCORE, na.rm = TRUE), 2),
    Skewness = round(skewness(RISK_SCORE, na.rm = TRUE), 2)
  ) %>%  
  select(region,N, Mean, Median, SD, Skewness)  # arrange columns in the desired order

# Print the summary table
gt_table <- gt(summary_table,caption="Summary Statistics of Risk Score across Regions")
gt_table
```

<br>
**8. Limitations and Improvements (20 points)**
<br>
The research goal was studying whether individuals who are more directly exposed to the impacts of climate-related hazards—such as flooding, hurricanes, extreme heat, or other forms of environmental disruption—may be more likely to perceive climate change as an urgent problem. The linear model examined the relationship between FEMA’s National Risk Index (RISK_SCORE)—the explanatory variable—and a measure of climate concern. Even though there is statistically positive correlation between two variables, we cannot conclude the climate risk leads to directly perceive climate change as an urgent problem. According to the FEMA regional data, there is a possibility that spacial autocorrelation between region could be generated. For improving the model, we have to conduct multiple regression considering region as a control variable.In addition, this research did not consider the possibility of temporal autocorrelation.For accurate research design, we can do use time series data or using year variable as an control variable for multiple regression. 

