---
title: 'Tutorial 8: Simple Linear Regression'
author: "Jiyeong Pyo"
editor_options:
  markdown:
    wrap: 72
output: html_document
---
*if beta =0 (null hypothesis), beta =!0: alternative 
t=beta hat =0 / standard error of beta hat 1
This tutorial will walk you through the steps of running a simple linear
regression (SLR) in R. I encourage you to read about Yale's
Environmental Performance Index (EPI) here:
<https://epi.envirocenter.yale.edu/>.

### Variables: GDP and Environmental Performance Index (EPI)

For the tutorial we will explore the relationship between GDP per capita
(GDPpc) and the Yale Environmental Performance Index.

The Environmental Performance Index (EPI.new) is an index for each
nation on its environmental performance. It is an index combined of
numerous weighted measures. Read more about the data on the Yale website
and the technical appendix 
https://epi.yale.edu/downloads/2024epitechnicalappendix20241207.pdf


```{r libraries,include=FALSE}
library(ggplot2)
library(dplyr)
library(WDI)
library(readr)
library(broom)
library(ggthemes)
```

We will read in the data with read_csv() (a readr function).

```{r import,include=FALSE}
rm(list=ls()) # removing objects
epi.df<-read_csv("~/EDE_Fall2025/Justina/ENV710/epi2024results.csv")
head(epi.df)
```

### Import GDP per capita data from World Bank (in current USD)

```{r import GDPpc,include=FALSE}
wb_data <-
  WDI(
    indicator = c("NY.GDP.PCAP.CD"), 
    country = "all", 
    start = 2024, 
    end = 2024,
    extra = TRUE
  )%>%
  glimpse()

```



```{r filter_only_country,include=FALSE}
wb_data <- wb_data %>%
  filter(region != "Aggregates") %>% #remove aggregates
  glimpse()
```

### Renaming the GDP variable

We will now rename "NY.GDP.PCAP.CD" to gdp_pc. The renamed variable will
be in the wb_data data frame.

```{r rename,include=FALSE}
wb_data <- rename(wb_data, gdp_pc = NY.GDP.PCAP.CD)
glimpse(wb_data)
```

Let's look at the gdp_pc data to make sure it looks good.

```{r head,include=FALSE}
head(wb_data, 10)
```

### Selecting variables that we will keep

Using the head and glimpse functions above, you saw that there are some
variables that we probably don't need. Let's clean that up by using the
**select()** function in dplyr. We will select the variables we want
using select() and then will look at the data frame with the glimpse()
function.

```{r select,include=FALSE}
wb_data<-wb_data%>%
  dplyr::select(iso3c, year, gdp_pc)%>%
  glimpse()
```

You should now have three variables in the data frame: iso3c, year, and
gdp_pc.

### Joining data Using the full_join() function in dplyr

We should now join wb_data with the epi.df using the three letter
country code that is in both data frames epi.df (iso) and wb_data
(iso3c). A full join joins all rows that are in either x or y. Call the
new joined data frame epi_gdp.df.

Read more about the joining functions in dplyr here:
<https://dplyr.tidyverse.org/reference/mutate-joins.html>
<https://statisticsglobe.com/r-dplyr-join-inner-left-right-full-semi-anti>

```{r join,include=FALSE}
epi_gdp.df <- full_join(epi.df, wb_data, by = c("iso" = "iso3c"))
```

We can now make our scatterplot.

### Scatterplots

You will now make a scatter plot of GDP per capita (x) and EPI.new (Y).

```{r scatter.plot,include=FALSE}

options(scipen = 999) #this bit of code turns off scientific notation

p1 <- epi_gdp.df %>% 
  ggplot(aes(x = gdp_pc, y = EPI.new, label = iso))+
       geom_point(size = 1.5) + 
       theme_minimal()+
       ylab("Environmental Performance Index")+
       xlab("Per Capita Gross Domestic Product (USD)")+
       ylim(0,100)+
       geom_text(aes(label = iso), size = 3, vjust=-0.5) #add text labels, vjust=vertical justification of labels
p1


```

You can see in the scatterplot that GDP per capita is positively skewed.
We should take the log of GDP per capita and then make the scatterplot
with the logged gdp_pc data. Use the mutate() function to add in
log_gdp_pc to the epi_gdp.df data frame. Do so in the chunk below.

```{r log_gdp,include=FALSE}
epi_gdp.df <- epi_gdp.df %>%
  mutate(log_gdp_pc = log(gdp_pc))

```

Develop a scatter plot of EPI.new \~ log_gdp_pc. Notice the units on the
Y-axis. You can set the limits of the y axis by adding the code:
ylim(0,100). Use geom_point to add text labels to the plot and set label=iso in the
aesthetic. Please add appropriate labels for the x- and y-axes.

```{r log_scatter,include=FALSE}
options(scipen = 999) #this bit of code turns off scientific notation

p2 <- epi_gdp.df %>% 
  ggplot(aes(x = log_gdp_pc, y = EPI.new, label = iso))+
       geom_point(size = 1.5) + 
       theme_minimal()+
       ylab("Environmental Performance Index")+
       xlab("Log of Per Capita Gross Domestic Product (USD)")+
       ylim(0,100)+
       geom_text(aes(label = iso), size = 3, vjust=-0.5) #add text labels
p2

```

We could also visualize the original data on a log scale of the x-axis
by adding scale_x_log10(). This adjusts the scale to log10. 


**Q1: What is the advantage of using this approach to visualizing the data vs logging the data and plotting the logged data? Any disadvantages?** The logged data helps to eliminating the skewness so it becomes more linear than the original data. 

```{r log10,include=FALSE}
p1+scale_x_log10() + 
  xlab("Per Capita Gross Domestic Product (USD)")
```

## Calculating correlations

The correlation coefficient (r) is a measure of linear relationship and
can range between -1 and 1.

Please calculate the correlation coefficients on both the original data
(EPI.new and gdp_pc) and the (natural) logged data of GDPpc.

The function for correlation coefficient is cor(). For correlations, it
does not matter which variable you place first. if there are missing
data, you should use the argument use="pairwise.complete.obs". Please
type ?cor in the console so you can see the arguments before running the
chunk below.

```{r correlation,include=FALSE}
#correlation coefficients for the original data
cor_original <- cor(epi_gdp.df$EPI.new,
                    epi_gdp.df$gdp_pc,
                    use = "pairwise.complete.obs")

cor_original

#correlation coefficients for the logged data of GDPpc
cor_log <- cor(epi_gdp.df$EPI.new,
                  epi_gdp.df$log_gdp_pc,
                  use = "pairwise.complete.obs")
cor_log
```



**Question 2**: What are the two correlation coefficients? Please interpret these
coefficients. The corelation coefficents means how both EPI and GDP per captia moves together. So it means, in original data, when 1 dollor of GDP per captia increased, EPI increased 0.7188428. In terms of the logged data, when 1 dollor of GDP per captia increased, EPI increased 0.7739801. 

## Simple linear model

The function to run a linear model in R is lm()

To see the arguments of lm() please type ?lm in the console.

The model is written Y~X.

```{r linearmodel}
model.1 <- lm(EPI.new ~ gdp_pc , data = epi_gdp.df)
model.1
```

The output of R gives us the equation for the simple linear model. The
equation can be written as **Predicted EPI.new = 41.00 + 0.0003437 * GDPpc** 


**Question 3: Why do you think the slope is such a small number?**

We can interpret this equation as, a one dollar increase in per capita
GDP is associated with a 0.0003437 increase in predicted Environmental
Performance Index. A dollar is such a small amount, so it is probably
better to interpret in terms of $1000 increase in GDPpc is associated
with a 0.3437 increase in predicted Environmental Performance Index. I calculated
this number by multiplying 1000 * 0.0003437.

We can also summarize the model to see the estimates, SE, t, and
p-values as well as the R-squared value and F-statistic.

```{r summarylm}
summary(model.1)
```

The test of significance (a t-test) on the coefficient tests whether the slope of
the line (of the populations) is zero. There is conclusive evidence to
suggest that the slope is different than zero (a flat line).

R-squared is the percent of the variation in the Y variable that is explained
by the X variable. In this case, 51.39%. We won't worry about adjusted R-squared
and the F statistic until we get to multiple linear regression.

We can now add a linear regression line to the plot. This is the line
that is drawn by ordinary least squares (OLS). We can use the function
geom_smooth and the argument method=lm to tell it to draw the line that
we calculated from a linear model.

This will draw the confidence interval around the line based on the
linear model.

```{r scatter.lm}
p3<-ggplot(data=epi_gdp.df, aes(x = gdp_pc, y = EPI.new)) +
  geom_point() + 
  geom_smooth(method=lm)+
  labs(x="GDP per capita (USD)", y="Environmental Performance Index")

p3
```

**Question 4: Describe the relations between EPI and GPDpc. Is this a linear relationship?**
It shows the GDP per capita is positivly related with environmental performance index. It shows linear relationship.


We should run the simple linear model using the log_gdp_pc data.

```{r,message=FALSE}
model.2 <- lm(EPI.new ~ log_gdp_pc , data = epi_gdp.df)
model.2
summary(model.2)
```

As you might have guessed, we now have a new equation when we use
logGDPpc. We no longer have a linear model, but rather what we call a
linear-log model. To interpret a linear-log model, an increase in x
(GDPpc) by one percent is associated with an increase in predicted y by
($\beta$/100) units (EPI).

So, a 1% increase in per capita GDP is associated with a 0.0631 increase
in predicted EPI. The R-squared is now 0.60.

Let's use the package broom to make a nice tidy tibble of the regression
results. First I'll turn scientific notation back on.

```{r}
options(scipen=0) #turn scientific notation back on. I'm turning it back on to get reasonable looking p-values
tidy(summary(model.2))
```

Let's make a new scatterplot with the loggedGDPpc data that includes the
regression line. Please run the chunk below.

```{r scatterlog, include=TRUE,message=FALSE}

p4<-ggplot(data=epi_gdp.df, aes(x = log_gdp_pc, y = EPI.new)) +
  geom_point() + 
  geom_smooth(method=lm)+
  labs(x="Logged GDP per capita (USD)", y="Environmental Performance Index")
p4
```

Awesome! That scatter plot looks pretty darn good. If we are meeting the
assumptions of ordinary least squares, we want equal spread of the
observations above and below the regression line across all values of
our explanatory variable (log gdp per capita).

**Question 5 (a, b, and c)**

**Now it is your turn to play around with one of the other variables in the Yale EPI dataset. 
Please look at its relationship with log GDP per capita (x). 

Please develop: <br>

(a) a scatterplot, <br>
```{r,message=FALSE, echo=FALSE}
p1 <- epi_gdp.df %>% 
  ggplot(aes(x = log_gdp_pc, y = OZD.new, label = iso))+
       geom_point(size = 1.5) + 
       theme_minimal()+
       ylab("Ozone Exposure")+
       xlab("Per Capita Gross Domestic Product (USD)")+
       ylim(0,100)+
       geom_text(aes(label = iso), size = 3, vjust=-0.5) 
p1

p1+scale_x_log10() + 
  xlab("Per Capita Gross Domestic Product (USD)")

```

(b) calculate a correlation coefficient, and  <br>
```{r,message=FALSE, echo=FALSE}
cor_log <- cor(epi_gdp.df$OZD.new,
                  epi_gdp.df$log_gdp_pc,
                  use = "pairwise.complete.obs")
cor_log
```

(c) Develop and interpret an appropriate simple linear regression model.** 
The logged GDP and the Ozone Exposure level has significant correlation (The Pr(>|t|)is 6.65e-09 *** so we can reject the null hypothesis.) 
```{r,message=FALSE, echo=FALSE}
p2<-ggplot(data=epi_gdp.df, aes(x = log_gdp_pc, y = OZD.new)) +
  geom_point() + 
  geom_smooth(method=lm)+
  labs(x="Logged GDP per capita (USD)", y="Ozone Exposure")
p2
```

```{r,message=FALSE, echo=FALSE}
model.2 <- lm(OZD.new ~ log_gdp_pc , data = epi_gdp.df)
model.2
summary(model.2)
```
```{r,message=FALSE}
options(scipen=0) 
tidy(summary(model.2))
```

Prior to selecting your variable, please skim the Policymaker's Summary, the EPI Report, and the Technical Appendix
which can be found here: https://epi.yale.edu/downloads 

**Please knit to html and upload to Assignments before lecture on Wednesday.**

### Congratulations on finishing Tutorial 8!

